### 2012-01-17 contents.tex
Transforming in short paper.

@@

Software development is a complex activity that demands the coordination of many people, such as programmers, designers, and managers. Software development activities should be coordinated in order to deliver products with high quality and on schedule. Many practices and processes have been proposed in order to aid the software development, targeting at high quality, low cost, fast delivery, or a combination of these.

Several researchers are concerned with the validity of studies based on bug tracking systems \cite{Aranda2009,Bird2009,Nguyen2010,Ayari2007}. The main concern is the accuracy of the information that is reported in such systems.
\end{section}

@@

Bug tracking systems are applications that allow users and developers of a software system to manage a list of bugs for the system. Usually, users and developers can report bugs, along with information such as steps to reproduce the bug and the operating system that was used. Then, developers choose bugs to fix and can report on the progress of the bug fixing activities, ask for clarification etc.

One important feature of a bug that is recorded on bug tracking systems is its status. The status records the progress of the bug fixing activity, and, as such, provides data to software engineering studies. Figure \ref{fig:bugzilla} shows each status that can be recorded in Bugzilla, a popular bug tracking system, along with typical transitions between statuses, i.e., the workflow. Other systems present comparable status and transitions.

@@

In the simplest cases, a bug is created and receive the status \texttt{UNCONFIRMED} or \texttt{NEW}. Next, it is assigned to a developer, and then it is resolved, possibly by fixing it with a patch on the source code. The solution is then verified by someone in the quality assurance team. If it passes the quality requirements, it is closed. If it doesn’t pass the quality requirements or if the solution only partially fixes the problem, the bug is reopened.
\end{subsection}

@@

@@

For the remaining 34 subprojects, we have selected only bugs that were verified until December 2009. The rationale is that if, within six months, a verified bug was not reopened, it probably will not be reopened after that and it is, thus, fully fixed.
\end{section}

>>>>

% http://wiki.eclipse.org/Bug_Reporting_FAQ
%   when a test pass comes up a component owner can build a query that searched for resolved bugs
%   After testing that the fix worked a resolved bug can be transitioned to verified, directly to closed, or in fact, reopened. 
%   By searching for bugs with a Status of verified and a resolution of fixed developers can come up with release notes.
%   Severity is assigned by a user and describes the level of impact the bug is having on them. Priority is assigned by the developer and describes the importance a developer places on fixing the bug.

@@

Bug repositories (or bug tracking systems) have since a long time been used in software projects to support coordination among stakeholders. Such systems record discussion and progress of software evolution activities, such as corrective, perfective, and ... changes \cite{Antoniol2008}. Hence, bug repositories are an opportunity to researchers who intend to investigate issues related to the quality of the product and of the process of a software development team, which can give insights about the nature of software development itself.

The information contained in bug repositories have been used in order to predict fault-proneness, to unveil developers' roles from data, to characterize transfer of work in software projects, and so on. Also, it has been used to investigate beliefs in software engineering related to software quality, such as the impact of developer turnover and of collective ownership on the occurrence of bugs.

However, mining bug repositories has its own risks. Previous research has identified problems of missing data (e.g., rationale, traceability links between reported bug fixes and source code changes) and inaccurate data (e.g., misclassification of bugs) \cite{Aranda2009,Bird2009,Nguyen2010,Ayari2007}.

\TODO why the VERIFIED status is so important?

In summary, for researchers who intend to use data from bug tracking systems, it is not sufficient to understand the data fields in a bug report the way they are documented in the system's documentation. The research should also investigate whether the data fields are used often, whether they are used accordingly to the documentation, and even if the data fields are used consistently across the developers.

In this paper, we investigate the status of a bug, a data field present in most bug tracking systems that reports the progress of the bug solving activity. In particular, we focus on the activity of verification of a fix. We are interested in answering the following exploratory research questions:

\begin{itemize}
	\item What activity is actually performed when a bug fix is marked as verified?
	\item When is the verification performed (i.e., just after the fix, or is there a specific period for the verification of fixed bugs?)
	\item How is the distribution of verification across developers?
	\item Is there a dedicated team for QA?
\end{itemize}

For research question xx, qualitative of a small sample. 
For research question xx, all bugs were considered.

@@

% Reposit ́orios de bugs. Arquivam relat ́orios de defeitos e solicitac ̧ ̃oes de mudan ̧ca. Tais reposito ́rios ajudam os desenvolvedores a registrar, discutir e acompanhar a situa ̧ca ̃o de cada solicita ̧c ̃ao, desde a atribui ̧ca ̃o de uma solicita ̧ca ̃o a um desenvolvedor at ́e a verifica ̧c ̃ao da corre ̧c ̃ao ou melhoria. Exemplos incluem o Bugzilla, o Trac e o Jira.

Bug tracking systems allow users and developers of a software system to manage a list of bugs for the system. Usually, users and developers can report bugs, along with information such as steps to reproduce the bug, its severity, the operating system used etc. Developers choose bugs to fix and can report on the progress of the bug fixing activities, ask for clarification etc.

One important feature of a bug that is recorded on bug tracking systems is its status. The status records the progress of the bug fixing activity, and, as such, provides data to software engineering studies. Figure \ref{fig:bugzilla} shows each status that can be recorded in Bugzilla, a popular bug tracking system, along with typical transitions between status values, i.e., the workflow. Other systems present comparable status values and transitions.

@@

In the simplest cases, a bug is created and receive the status \texttt{UNCONFIRMED} (if it was created by a regular user) or \texttt{NEW} (if it was created by a developer). Next, it is assigned to a developer, and then it is resolved, possibly by fixing it with a patch on the source code. The solution is then verified by someone in the quality assurance team. If it passes the quality requirements, it is closed and the patch is added to the next release. If it doesn’t pass the quality requirements or if the solution only partially fixes the problem, the bug is reopened.

\end{subsection}

\begin{section}{Method}

\begin{section}{Data Extraction}

	For the remaining 34 subprojects, we have selected only bugs that were verified until December 2009. The rationale is that if, within six months, a verified bug was not reopened, it probably will not be reopened after that and it is, thus, fully fixed.

@@

\begin{section}{Data Sampling}
	Scripts to classify interesting cases.
	Random sampling stratified by case.

\begin{section}{Analysis}
	grounded theory-style.

\begin{section}{Conclusion}

@@
